---
title: "Morph_v_DNA"
author: "Savannah Ryburn"
date: "2025-05-20"
output: html_document
---

```{r}
library(tidyverse)
library(lubridate)
library(dplyr)
library(tidyr)
library(cowplot)
library(scales)
library(ggplot2)
library(phyloseq)
library(RColorBrewer)
```

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "/Users/savannahryburn/Desktop/Stomach_Chunk_data/GitHub_Morph_v_DNA/GitHub_Morph_v_DNA")

Chunk <- read.csv("Data/Stomach_Chunk_Data.csv")
Chunk_simple <- read.csv("Data/Chunk_simple.csv")
Richness_diff_data <- read.csv("Data/Richness_SharkSpecies_Diff.csv")
Chunk_per <- read.csv("Data/Chunk_data_permanova.csv")
```

#Comparing taxonomic rank
```{r}
species_visual <- ggplot(Chunk, 
       aes(x = factor(Shark_species),
           fill = factor(lowest_visual, 
                         levels = c("Subphylum", "Class", "Order", "Family", "Genus", "Species", "NA"),
                         labels = c("Subphylum", "Class", "Order", "Family", "Genus", "Species", "NA")))) + 
  geom_bar(position = "fill") +
  scale_y_continuous(breaks = seq(0, 1, .2), 
                     label = percent, expand = c(0,0)) +
  labs(y = "Percent of prey taxa identified", 
       fill = "Prey taxa",
       x = "Shark species") +
  theme_classic() +
    scale_fill_manual(values = c("indianred4", "royalblue", "aquamarine4", "darkolivegreen3", "khaki", "lightblue", "grey")) +
   theme(text = element_text(size =11))

 species_visual

library(scales)
species_DNA <- ggplot(Chunk, 
       aes(x = factor(Shark_species),
           fill = factor(BLAST_taxanomic_resolution, 
                         levels = c("Order", "Family", "Genus", "Species", "NA"),
                         labels = c("Order", "Family", "Genus", "Species", "NA")))) + 
  geom_bar(position = "fill") +
  scale_y_continuous(breaks = seq(0, 1, .2), 
                     label = percent, expand = c(0,0)) +
  labs(y = "Percent of prey taxa identified", 
       fill = "Prey taxa",
       x = "Shark species") +
  theme_classic() +
  scale_fill_manual(values = c("aquamarine4", "darkolivegreen3", "khaki", "lightblue", "grey")) +
   theme(text = element_text(size =11))

species_DNA


plot_grid(species_visual, species_DNA, labels = "AUTO")
```

Make Percentage Tables
```{r}
####Visual####
# Calculate percentages
percentage_table <- Chunk %>%
  group_by(Shark_species, lowest_visual) %>%
  summarise(count = n(), .groups = "drop") %>%
  group_by(Shark_species) %>%
  mutate(percent = count / sum(count) * 100) %>%
  ungroup()

# Pivot into a wide table for better readability
wide_table <- percentage_table %>%
  select(Shark_species, lowest_visual, percent) %>%
  pivot_wider(names_from = lowest_visual, values_from = percent, values_fill = 0)

# View the table
print(wide_table)

# Optional: Save the table as a CSV
#write.csv(wide_table, "percentage_table.csv", row.names = FALSE)


####DNA####
# Calculate percentages
dna_percentage_table <- Chunk %>%
  group_by(Shark_species, BLAST_taxanomic_resolution) %>%
  summarise(count = n(), .groups = "drop") %>%
  group_by(Shark_species) %>%
  mutate(percent = count / sum(count) * 100) %>%
  ungroup()

# Pivot into a wide table for better readability
dna_wide_table <- dna_percentage_table %>%
  select(Shark_species, BLAST_taxanomic_resolution, percent) %>%
  pivot_wider(names_from = BLAST_taxanomic_resolution, values_from = percent, values_fill = 0)

# View the table
print(dna_wide_table)

# Optional: Save the table as a CSV
#write.csv(dna_wide_table, "dna_percentage_table.csv", row.names = FALSE)
```

Chi Squared
```{r}
# Step 1: Create datasets for each methodology
visual_data <- Chunk_simple %>% 
  select(lowest_visual) %>% 
  mutate(Methodology = "Visual", Count = 1) %>%
  rename(Value = lowest_visual)

dna_data <- Chunk_simple %>% 
  select(BLAST_taxanomic_resolution) %>% 
  mutate(Methodology = "DNA", Count = 1) %>%
  rename(Value = BLAST_taxanomic_resolution)

# Step 2: Combine the datasets
combined_data <- bind_rows(visual_data, dna_data)
View(combined_data)

# Step 3: Plot combined data on the same x-axis (not using this plot anymore)
# Count_method_combined_plot <- ggplot(combined_data, aes(x = Methodology, fill = Value)) +
#   geom_bar(position = "dodge") +  # Bars side-by-side
#   labs(x = "Methodology", y = "Count of Prey Items", fill = "Taxonomic resolution") +
#   scale_fill_manual(values = c("navy", "yellowgreen")) +
#   theme_classic()
# 
# Count_method_combined_plot


#Chi-squared test
# Step 1: Create a contingency table
contingency_table <- table(combined_data$Methodology, combined_data$Value)
View(contingency_table) ##used this table instead of the above plot

# Step 2: Perform chi-squared test
chi_squared_test <- chisq.test(contingency_table)

# Display results
chi_squared_test
```

#Count match
```{r}
Chunk$Accuracy <- factor(Chunk$Accuracy, levels = c("Match", "General", "Unidentified general", "Too specific", "Wrong visual", "NA DNA"))

Count_method_plot <- ggplot(Chunk, aes(x=Accuracy,fill=Shark_species)) + geom_bar() + labs(x ="Accuracy", y = "Count of Prey Items") + theme_classic() + scale_fill_brewer(palette = "GnBu")

Count_method_plot


#Fisher Test
contingency_table <- table(Chunk$Accuracy, Chunk$Shark_species)
fisher.test(contingency_table, simulate.p.value = TRUE, B = 1e6)
```

#Figure out what to compare Accuracy to here (BLAST_ID)
```{r}
Chunk$Accuracy <- factor(Chunk$Accuracy, levels = c("Match", "General", "Unidentified general", "Too specific", "Wrong visual", "NA DNA"))

Count_method_plot <- ggplot(Chunk, aes(x=Accuracy,fill=BLAST_ID)) + geom_bar() + labs(x ="Accuracy", y = "Count of Prey Items") + theme_classic()

Count_method_plot


table_summary <- table(Chunk$BLAST_ID, Chunk$Accuracy)
table_summary


#Fisher Test
contingency_table <- table(Chunk$Accuracy, Chunk$BLAST_ID)
fisher.test(contingency_table, simulate.p.value = TRUE, B = 1e6)
```

#Figure out what to compare Accuracy to here (Visual_ID)
```{r}
Chunk$Accuracy <- factor(Chunk$Accuracy, levels = c("Match", "General", "Unidentified general", "Too specific", "Wrong visual", "NA DNA"))

Count_method_plot <- ggplot(Chunk, aes(x=Accuracy,fill=Visual_ID)) + geom_bar() + labs(x ="Accuracy", y = "Count of Prey Items") + theme_classic()

Count_method_plot


table_summary <- table(Chunk$Visual_ID, Chunk$Accuracy)
table_summary


#Fisher Test
contingency_table <- table(Chunk$Accuracy, Chunk$Visual_ID)
fisher.test(contingency_table, simulate.p.value = TRUE, B = 1e6)
```

Match vs. prey weight
```{r}
ggplot(Chunk, aes(x = Accuracy, y = Prey_Weight)) +
  geom_boxplot(outlier.shape = NA) + # Suppress outliers from box plot
  geom_jitter(width = 0.2, alpha = 0.5, color = "black") +
  labs(x = "Accuracy", y = "Prey Weight") +
  theme_classic()
```

IRI% Calculations - All sharks combined
```{r}
###Visual

# Remove rows with NA values in Count or Weight
Chunk <- Chunk %>%
  filter(!is.na(Number.of.individuals.in.sample), !is.na(Prey_Weight))
Chunk

# Check for zero total counts or weights
sum(Chunk$Number.of.individuals.in.sample)
sum(Chunk$Prey_Weight)


# Frequency of Occurrence (FO) calculation
fo <- Chunk %>%
  group_by(Visual_ID) %>%
  summarise(Frequency = n_distinct(Fodrie_shark)) %>%
  mutate(FO = (Frequency / n_distinct(Chunk$Fodrie_shark)) * 100)

# Numerical Abundance (N%) calculation
n_percentage <- Chunk %>%
  group_by(Visual_ID) %>%
  summarise(Total_Count = sum(Number.of.individuals.in.sample)) %>%
  mutate(N_percent = (Total_Count / sum(Total_Count)) * 100)

# Weight Percentage (W%) calculation
w_percentage <- Chunk %>%
  group_by(Visual_ID) %>%
  summarise(Total_Weight = sum(Prey_Weight)) %>%
  mutate(W_percent = (Total_Weight / sum(Total_Weight)) * 100)

# Combine and calculate IRI
iri_data_V <- fo %>%
  left_join(n_percentage, by = "Visual_ID") %>%
  left_join(w_percentage, by = "Visual_ID") %>%
  mutate(IRI = (N_percent + W_percent) * FO)

iri_data_V

# Calculate the IRI percentage
total_iri <- sum(iri_data_V$IRI, na.rm = TRUE)
iri_data_V <- iri_data_V %>%
  mutate(IRI_percent = (IRI / total_iri) * 100)

iri_data_V

#write.csv(iri_data_V, "/Users/savannahryburn/Desktop/iri_data_Visual.csv", row.names = FALSE)

###DNA
# Frequency of Occurrence (FO) calculation
fo <- Chunk %>%
  group_by(BLAST_ID) %>%
  summarise(Frequency = n_distinct(Fodrie_shark)) %>%
  mutate(FO = (Frequency / n_distinct(Chunk$Fodrie_shark)) * 100)

# Numerical Abundance (N%) calculation
n_percentage <- Chunk %>%
  group_by(BLAST_ID) %>%
  summarise(Total_Count = sum(Number.of.individuals.in.sample)) %>%
  mutate(N_percent = (Total_Count / sum(Total_Count)) * 100)

# Weight Percentage (W%) calculation
w_percentage <- Chunk %>%
  group_by(BLAST_ID) %>%
  summarise(Total_Weight = sum(Prey_Weight)) %>%
  mutate(W_percent = (Total_Weight / sum(Total_Weight)) * 100)

# Combine and calculate IRI
iri_data_DNA <- fo %>%
  left_join(n_percentage, by = "BLAST_ID") %>%
  left_join(w_percentage, by = "BLAST_ID") %>%
  mutate(IRI = (N_percent + W_percent) * FO)

iri_data_DNA

# Calculate the IRI percentage
total_iri <- sum(iri_data_DNA$IRI, na.rm = TRUE)
iri_data_DNA <- iri_data_DNA %>%
  mutate(IRI_percent = (IRI / total_iri) * 100)

iri_data_DNA

#write.csv(iri_data_DNA, "/Users/savannahryburn/Desktop/iri_data_DNA.csv", row.names = FALSE)
```

IRI% Calculations - Atlantic Sharpnose
```{r}
###Visual

# Remove rows with NA values in Count or Weight
Chunk <- Chunk %>%
  filter(!is.na(Number.of.individuals.in.sample), !is.na(Prey_Weight))
Chunk

# Check for zero total counts or weights
sum(Chunk$Number.of.individuals.in.sample)
sum(Chunk$Prey_Weight)

#Now only by shark species (Atlantic sharpnose)
Chunk_AS <- Chunk %>%
  filter(Shark_species == "Atlantic Sharpnose")
View(Chunk_AS)

# Frequency of Occurrence (FO) calculation
fo <- Chunk_AS %>%
  group_by(Visual_ID) %>%
  summarise(Frequency = n_distinct(Fodrie_shark)) %>%
  mutate(FO = (Frequency / n_distinct(Chunk_AS$Fodrie_shark)) * 100)

# Numerical Abundance (N%) calculation
n_percentage <- Chunk_AS %>%
  group_by(Visual_ID) %>%
  summarise(Total_Count = sum(Number.of.individuals.in.sample)) %>%
  mutate(N_percent = (Total_Count / sum(Total_Count)) * 100)

# Weight Percentage (W%) calculation
w_percentage <- Chunk_AS %>%
  group_by(Visual_ID) %>%
  summarise(Total_Weight = sum(Prey_Weight)) %>%
  mutate(W_percent = (Total_Weight / sum(Total_Weight)) * 100)

# Combine and calculate IRI
iri_data_V_AS <- fo %>%
  left_join(n_percentage, by = "Visual_ID") %>%
  left_join(w_percentage, by = "Visual_ID") %>%
  mutate(IRI = (N_percent + W_percent) * FO)

iri_data_V_AS

# Calculate the IRI percentage
total_iri <- sum(iri_data_V_AS$IRI, na.rm = TRUE)
iri_data_V_AS <- iri_data_V_AS %>%
  mutate(IRI_percent = (IRI / total_iri) * 100)

iri_data_V_AS

#write.csv(iri_data_V_AS, "/Users/savannahryburn/Desktop/Stomach_Chunk_data/iri_data_Visual_AS.csv", row.names = FALSE)

###DNA
# Frequency of Occurrence (FO) calculation
fo <- Chunk_AS %>%
  group_by(BLAST_ID) %>%
  summarise(Frequency = n_distinct(Fodrie_shark)) %>%
  mutate(FO = (Frequency / n_distinct(Chunk_AS$Fodrie_shark)) * 100)

# Numerical Abundance (N%) calculation
n_percentage <- Chunk_AS %>%
  group_by(BLAST_ID) %>%
  summarise(Total_Count = sum(Number.of.individuals.in.sample)) %>%
  mutate(N_percent = (Total_Count / sum(Total_Count)) * 100)

# Weight Percentage (W%) calculation
w_percentage <- Chunk_AS %>%
  group_by(BLAST_ID) %>%
  summarise(Total_Weight = sum(Prey_Weight)) %>%
  mutate(W_percent = (Total_Weight / sum(Total_Weight)) * 100)

# Combine and calculate IRI
iri_data_DNA_AS <- fo %>%
  left_join(n_percentage, by = "BLAST_ID") %>%
  left_join(w_percentage, by = "BLAST_ID") %>%
  mutate(IRI = (N_percent + W_percent) * FO)

iri_data_DNA_AS

# Calculate the IRI percentage
total_iri <- sum(iri_data_DNA_AS$IRI, na.rm = TRUE)
iri_data_DNA_AS <- iri_data_DNA_AS %>%
  mutate(IRI_percent = (IRI / total_iri) * 100)

iri_data_DNA_AS

#write.csv(iri_data_DNA_AS, "/Users/savannahryburn/Desktop/Stomach_Chunk_data/iri_data_DNA_AS.csv", row.names = FALSE)
```

IRI% Calculations - Blacknose
```{r}
###Visual

# Remove rows with NA values in Count or Weight
Chunk <- Chunk %>%
  filter(!is.na(Number.of.individuals.in.sample), !is.na(Prey_Weight))
Chunk

# Check for zero total counts or weights
sum(Chunk$Number.of.individuals.in.sample)
sum(Chunk$Prey_Weight)

#Now only by shark species (Atlantic sharpnose)
Chunk_BN <- Chunk %>%
  filter(Shark_species == "Blacknose")
#View(Chunk_BN)

# Frequency of Occurrence (FO) calculation
fo <- Chunk_BN %>%
  group_by(Visual_ID) %>%
  summarise(Frequency = n_distinct(Fodrie_shark)) %>%
  mutate(FO = (Frequency / n_distinct(Chunk_BN$Fodrie_shark)) * 100)

# Numerical Abundance (N%) calculation
n_percentage <- Chunk_BN %>%
  group_by(Visual_ID) %>%
  summarise(Total_Count = sum(Number.of.individuals.in.sample)) %>%
  mutate(N_percent = (Total_Count / sum(Total_Count)) * 100)

# Weight Percentage (W%) calculation
w_percentage <- Chunk_BN %>%
  group_by(Visual_ID) %>%
  summarise(Total_Weight = sum(Prey_Weight)) %>%
  mutate(W_percent = (Total_Weight / sum(Total_Weight)) * 100)

# Combine and calculate IRI
iri_data_V_BN <- fo %>%
  left_join(n_percentage, by = "Visual_ID") %>%
  left_join(w_percentage, by = "Visual_ID") %>%
  mutate(IRI = (N_percent + W_percent) * FO)

iri_data_V_BN

# Calculate the IRI percentage
total_iri <- sum(iri_data_V_BN$IRI, na.rm = TRUE)
iri_data_V_BN <- iri_data_V_BN %>%
  mutate(IRI_percent = (IRI / total_iri) * 100)

iri_data_V_BN

#write.csv(iri_data_V_BN, "/Users/savannahryburn/Desktop/Stomach_Chunk_data/iri_data_Visual_BN.csv", row.names = FALSE)

###DNA
# Frequency of Occurrence (FO) calculation
fo <- Chunk_BN %>%
  group_by(BLAST_ID) %>%
  summarise(Frequency = n_distinct(Fodrie_shark)) %>%
  mutate(FO = (Frequency / n_distinct(Chunk_BN$Fodrie_shark)) * 100)

# Numerical Abundance (N%) calculation
n_percentage <- Chunk_BN %>%
  group_by(BLAST_ID) %>%
  summarise(Total_Count = sum(Number.of.individuals.in.sample)) %>%
  mutate(N_percent = (Total_Count / sum(Total_Count)) * 100)

# Weight Percentage (W%) calculation
w_percentage <- Chunk_BN %>%
  group_by(BLAST_ID) %>%
  summarise(Total_Weight = sum(Prey_Weight)) %>%
  mutate(W_percent = (Total_Weight / sum(Total_Weight)) * 100)

# Combine and calculate IRI
iri_data_DNA_BN <- fo %>%
  left_join(n_percentage, by = "BLAST_ID") %>%
  left_join(w_percentage, by = "BLAST_ID") %>%
  mutate(IRI = (N_percent + W_percent) * FO)

iri_data_DNA_BN

# Calculate the IRI percentage
total_iri <- sum(iri_data_DNA_BN$IRI, na.rm = TRUE)
iri_data_DNA_BN <- iri_data_DNA_BN %>%
  mutate(IRI_percent = (IRI / total_iri) * 100)

iri_data_DNA_BN

#write.csv(iri_data_DNA_BN, "/Users/savannahryburn/Desktop/Stomach_Chunk_data/iri_data_DNA_BN.csv", row.names = FALSE)
```

IRI% Calculations - Blacktip
```{r}
###Visual

# Remove rows with NA values in Count or Weight
Chunk <- Chunk %>%
  filter(!is.na(Number.of.individuals.in.sample), !is.na(Prey_Weight))
Chunk

# Check for zero total counts or weights
sum(Chunk$Number.of.individuals.in.sample)
sum(Chunk$Prey_Weight)

#Now only by shark species (Atlantic sharpnose)
Chunk_BT <- Chunk %>%
  filter(Shark_species == "Blacktip")
#View(Chunk_BT)

# Frequency of Occurrence (FO) calculation
fo <- Chunk_BT %>%
  group_by(Visual_ID) %>%
  summarise(Frequency = n_distinct(Fodrie_shark)) %>%
  mutate(FO = (Frequency / n_distinct(Chunk_BT$Fodrie_shark)) * 100)

# Numerical Abundance (N%) calculation
n_percentage <- Chunk_BT %>%
  group_by(Visual_ID) %>%
  summarise(Total_Count = sum(Number.of.individuals.in.sample)) %>%
  mutate(N_percent = (Total_Count / sum(Total_Count)) * 100)

# Weight Percentage (W%) calculation
w_percentage <- Chunk_BT %>%
  group_by(Visual_ID) %>%
  summarise(Total_Weight = sum(Prey_Weight)) %>%
  mutate(W_percent = (Total_Weight / sum(Total_Weight)) * 100)

# Combine and calculate IRI
iri_data_V_BT <- fo %>%
  left_join(n_percentage, by = "Visual_ID") %>%
  left_join(w_percentage, by = "Visual_ID") %>%
  mutate(IRI = (N_percent + W_percent) * FO)

iri_data_V_BT

# Calculate the IRI percentage
total_iri <- sum(iri_data_V_BT$IRI, na.rm = TRUE)
iri_data_V_BT <- iri_data_V_BT %>%
  mutate(IRI_percent = (IRI / total_iri) * 100)

iri_data_V_BT

#write.csv(iri_data_V_BT, "/Users/savannahryburn/Desktop/Stomach_Chunk_data/iri_data_Visual_BT.csv", row.names = FALSE)

###DNA
# Frequency of Occurrence (FO) calculation
fo <- Chunk_BT %>%
  group_by(BLAST_ID) %>%
  summarise(Frequency = n_distinct(Fodrie_shark)) %>%
  mutate(FO = (Frequency / n_distinct(Chunk_BT$Fodrie_shark)) * 100)

# Numerical Abundance (N%) calculation
n_percentage <- Chunk_BT %>%
  group_by(BLAST_ID) %>%
  summarise(Total_Count = sum(Number.of.individuals.in.sample)) %>%
  mutate(N_percent = (Total_Count / sum(Total_Count)) * 100)

# Weight Percentage (W%) calculation
w_percentage <- Chunk_BT %>%
  group_by(BLAST_ID) %>%
  summarise(Total_Weight = sum(Prey_Weight)) %>%
  mutate(W_percent = (Total_Weight / sum(Total_Weight)) * 100)

# Combine and calculate IRI
iri_data_DNA_BT <- fo %>%
  left_join(n_percentage, by = "BLAST_ID") %>%
  left_join(w_percentage, by = "BLAST_ID") %>%
  mutate(IRI = (N_percent + W_percent) * FO)

iri_data_DNA_BT

# Calculate the IRI percentage
total_iri <- sum(iri_data_DNA_BT$IRI, na.rm = TRUE)
iri_data_DNA_BT <- iri_data_DNA_BT %>%
  mutate(IRI_percent = (IRI / total_iri) * 100)

iri_data_DNA_BT

#write.csv(iri_data_DNA_BT, "/Users/savannahryburn/Desktop/Stomach_Chunk_data/iri_data_DNA_BT.csv", row.names = FALSE)
```

IRI% Calculations - Bonnethead
```{r}
###Visual

# Remove rows with NA values in Count or Weight
Chunk <- Chunk %>%
  filter(!is.na(Number.of.individuals.in.sample), !is.na(Prey_Weight))
Chunk

# Check for zero total counts or weights
sum(Chunk$Number.of.individuals.in.sample)
sum(Chunk$Prey_Weight)

#Now only by shark species (Atlantic sharpnose)
Chunk_BH <- Chunk %>%
  filter(Shark_species == "Bonnethead")
#View(Chunk_BH)

# Frequency of Occurrence (FO) calculation
fo <- Chunk_BH %>%
  group_by(Visual_ID) %>%
  summarise(Frequency = n_distinct(Fodrie_shark)) %>%
  mutate(FO = (Frequency / n_distinct(Chunk_BH$Fodrie_shark)) * 100)

# Numerical Abundance (N%) calculation
n_percentage <- Chunk_BH %>%
  group_by(Visual_ID) %>%
  summarise(Total_Count = sum(Number.of.individuals.in.sample)) %>%
  mutate(N_percent = (Total_Count / sum(Total_Count)) * 100)

# Weight Percentage (W%) calculation
w_percentage <- Chunk_BH %>%
  group_by(Visual_ID) %>%
  summarise(Total_Weight = sum(Prey_Weight)) %>%
  mutate(W_percent = (Total_Weight / sum(Total_Weight)) * 100)

# Combine and calculate IRI
iri_data_V_BH <- fo %>%
  left_join(n_percentage, by = "Visual_ID") %>%
  left_join(w_percentage, by = "Visual_ID") %>%
  mutate(IRI = (N_percent + W_percent) * FO)

iri_data_V_BH

# Calculate the IRI percentage
total_iri <- sum(iri_data_V_BH$IRI, na.rm = TRUE)
iri_data_V_BH <- iri_data_V_BH %>%
  mutate(IRI_percent = (IRI / total_iri) * 100)

iri_data_V_BH

#write.csv(iri_data_V_BH, "/Users/savannahryburn/Desktop/Stomach_Chunk_data/iri_data_Visual_BH.csv", row.names = FALSE)

###DNA
# Frequency of Occurrence (FO) calculation
fo <- Chunk_BH %>%
  group_by(BLAST_ID) %>%
  summarise(Frequency = n_distinct(Fodrie_shark)) %>%
  mutate(FO = (Frequency / n_distinct(Chunk_BH$Fodrie_shark)) * 100)

# Numerical Abundance (N%) calculation
n_percentage <- Chunk_BH %>%
  group_by(BLAST_ID) %>%
  summarise(Total_Count = sum(Number.of.individuals.in.sample)) %>%
  mutate(N_percent = (Total_Count / sum(Total_Count)) * 100)

# Weight Percentage (W%) calculation
w_percentage <- Chunk_BH %>%
  group_by(BLAST_ID) %>%
  summarise(Total_Weight = sum(Prey_Weight)) %>%
  mutate(W_percent = (Total_Weight / sum(Total_Weight)) * 100)

# Combine and calculate IRI
iri_data_DNA_BH <- fo %>%
  left_join(n_percentage, by = "BLAST_ID") %>%
  left_join(w_percentage, by = "BLAST_ID") %>%
  mutate(IRI = (N_percent + W_percent) * FO)

iri_data_DNA_BH

# Calculate the IRI percentage
total_iri <- sum(iri_data_DNA_BH$IRI, na.rm = TRUE)
iri_data_DNA_BH <- iri_data_DNA_BH %>%
  mutate(IRI_percent = (IRI / total_iri) * 100)

iri_data_DNA_BH

#write.csv(iri_data_DNA_BH, "/Users/savannahryburn/Desktop/Stomach_Chunk_data/iri_data_DNA_BH.csv", row.names = FALSE)
```

#PERMANOVA comparing differences between shark species diet and method
```{r}
library(vegan)
library(tidyr)
library(dplyr)

# Convert BLAST_ID (DNA barcoding results) to presence-absence matrix
barcode_pa <- Chunk_per %>%
  filter(!is.na(BLAST_ID) & BLAST_ID != "") %>%  # Remove NA and empty species names
  mutate(Present = 1) %>%  # Create presence indicator
  pivot_wider(names_from = BLAST_ID, values_from = Present, values_fill = 0)

# Convert Visual_ID (Morphological results) to presence-absence matrix
morpho_pa <- Chunk_per %>%
  filter(!is.na(Visual_ID) & Visual_ID != "") %>%  # Remove NA and empty values
  mutate(Present = 1) %>%  
  pivot_wider(names_from = Visual_ID, values_from = Present, values_fill = 0)

# Remove non-numeric columns
barcode_matrix <- barcode_pa %>% select(where(is.numeric))
morpho_matrix <- morpho_pa %>% select(where(is.numeric))

# Compute Jaccard distance (for presence-absence data)
sum(is.na(barcode_matrix))
barcode_matrix[is.na(barcode_matrix)] <- 0
barcode_matrix <- na.omit(barcode_matrix)
barcode_dist <- vegdist(barcode_matrix, method = "jaccard")

sum(is.na(morpho_matrix))
morpho_matrix[is.na(morpho_matrix)] <- 0
morpho_matrix <- na.omit(morpho_matrix)
morpho_dist <- vegdist(morpho_matrix, method = "jaccard")

barcode_perm <- adonis2(barcode_dist ~ Shark_species, data = barcode_pa, permutations = 999)
morpho_perm <- adonis2(morpho_dist ~ Shark_species, data = morpho_pa, permutations = 999)

print(barcode_perm)
print(morpho_perm)



# Calculate dissimilarity matrices for both datasets
barcode_dist <- vegdist(barcode_matrix, method = "jaccard")  # Barcoding data
morpho_dist <- vegdist(morpho_matrix, method = "jaccard")    # Morphological data

# Convert dissimilarity matrices to data frames
barcode_df <- as.data.frame(as.matrix(barcode_dist))
morpho_df <- as.data.frame(as.matrix(morpho_dist))

# Add a 'method' column to distinguish between morphological and barcoding methods
barcode_df$method <- "Barcoding"
morpho_df$method <- "Morphological"

# Combine the two dataframes (one for barcoding and one for morphological data)
combined_df <- rbind(barcode_df, morpho_df)

# Add Shark_species information to the combined dataset
combined_df$Shark_Species <- rep(barcode_pa$Shark_species, 2)  # Assuming same samples for both methods

# Select only the numeric columns for the dissimilarity matrix
combined_df_numeric <- combined_df %>%
  select_if(is.numeric)

# Run PERMANOVA with method and Shark_Species as grouping factors
combined_perm <- adonis2(combined_df_numeric ~ method + Shark_Species, data = combined_df, permutations = 999)
print(combined_perm)
```

#PERMANOVA comparing differences between methods (separated by shark species)
```{r}
library(vegan)
library(tidyr)
library(dplyr)

####### Blacktip ########

#Now only by shark species (Blacktip) - this is FOO
# Filter the dataset for only Blacktip shark species
Chunk_per_BT <- Chunk_per %>% filter(Shark_species == "Blacktip")

# Convert BLAST_ID (DNA barcoding results) to presence-absence matrix
barcode_pa <- Chunk_per_BT %>%
  filter(!is.na(BLAST_ID) & BLAST_ID != "") %>%  # Remove NA and empty species names
  mutate(Present = 1) %>%  # Create presence indicator
  pivot_wider(names_from = BLAST_ID, values_from = Present, values_fill = 0)

# Convert Visual_ID (Morphological results) to presence-absence matrix
morpho_pa <- Chunk_per_BT %>%
  filter(!is.na(Visual_ID) & Visual_ID != "") %>%  # Remove NA and empty values
  mutate(Present = 1) %>%  
  pivot_wider(names_from = Visual_ID, values_from = Present, values_fill = 0)

# Remove non-numeric columns
barcode_matrix <- barcode_pa %>% select(where(is.numeric))
morpho_matrix <- morpho_pa %>% select(where(is.numeric))

# Calculate dissimilarity matrices for both datasets
barcode_dist <- vegdist(barcode_matrix, method = "jaccard")  # Barcoding data
morpho_dist <- vegdist(morpho_matrix, method = "jaccard")    # Morphological data

# Convert dissimilarity matrices to data frames
barcode_df <- as.data.frame(as.matrix(barcode_dist))
morpho_df <- as.data.frame(as.matrix(morpho_dist))

# Add a 'method' column to distinguish between morphological and barcoding methods
barcode_df$method <- "Barcoding"
morpho_df$method <- "Morphological"

# Combine the two dataframes (one for barcoding and one for morphological data)
combined_df <- rbind(barcode_df, morpho_df)

# Add Shark_species information to the combined dataset
combined_df$Shark_Species <- rep(barcode_pa$Shark_species, 2)  # Assuming same samples for both methods

# Select only the numeric columns for the dissimilarity matrix
combined_df_numeric <- combined_df %>%
  select_if(is.numeric)

# Run PERMANOVA with method and Shark_Species as grouping factors
combined_perm <- adonis2(combined_df_numeric ~ method, data = combined_df, permutations = 999)
print(combined_perm)

####### Blacknose ########

#Now only by shark species (Blacknose) - this is FOO
# Filter the dataset for only Blacknose shark species
Chunk_per_BN <- Chunk_per %>% filter(Shark_species == "Blacknose")

# Convert BLAST_ID (DNA barcoding results) to presence-absence matrix
barcode_pa <- Chunk_per_BN %>%
  filter(!is.na(BLAST_ID) & BLAST_ID != "") %>%  # Remove NA and empty species names
  mutate(Present = 1) %>%  # Create presence indicator
  pivot_wider(names_from = BLAST_ID, values_from = Present, values_fill = 0)

# Convert Visual_ID (Morphological results) to presence-absence matrix
morpho_pa <- Chunk_per_BN %>%
  filter(!is.na(Visual_ID) & Visual_ID != "") %>%  # Remove NA and empty values
  mutate(Present = 1) %>%  
  pivot_wider(names_from = Visual_ID, values_from = Present, values_fill = 0)

# Remove non-numeric columns
barcode_matrix <- barcode_pa %>% select(where(is.numeric))
morpho_matrix <- morpho_pa %>% select(where(is.numeric))

# Calculate dissimilarity matrices for both datasets
barcode_dist <- vegdist(barcode_matrix, method = "jaccard")  # Barcoding data
morpho_dist <- vegdist(morpho_matrix, method = "jaccard")    # Morphological data

# Convert dissimilarity matrices to data frames
barcode_df <- as.data.frame(as.matrix(barcode_dist))
morpho_df <- as.data.frame(as.matrix(morpho_dist))

# Add a 'method' column to distinguish between morphological and barcoding methods
barcode_df$method <- "Barcoding"
morpho_df$method <- "Morphological"

# Combine the two dataframes (one for barcoding and one for morphological data)
combined_df <- rbind(barcode_df, morpho_df)

# Add Shark_species information to the combined dataset
combined_df$Shark_Species <- rep(barcode_pa$Shark_species, 2)  # Assuming same samples for both methods

# Select only the numeric columns for the dissimilarity matrix
combined_df_numeric <- combined_df %>%
  select_if(is.numeric)

# Run PERMANOVA with method and Shark_Species as grouping factors
combined_perm <- adonis2(combined_df_numeric ~ method, data = combined_df, permutations = 999)
print(combined_perm)


####### Bonnethead ########

#Now only by shark species (Bonnethead) - this is FOO
# Filter the dataset for only Bonnethead shark species
Chunk_per_BH <- Chunk_per %>% filter(Shark_species == "Bonnethead")

# Convert BLAST_ID (DNA barcoding results) to presence-absence matrix
barcode_pa <- Chunk_per_BH %>%
  filter(!is.na(BLAST_ID) & BLAST_ID != "") %>%  # Remove NA and empty species names
  mutate(Present = 1) %>%  # Create presence indicator
  pivot_wider(names_from = BLAST_ID, values_from = Present, values_fill = 0)

# Convert Visual_ID (Morphological results) to presence-absence matrix
morpho_pa <- Chunk_per_BH %>%
  filter(!is.na(Visual_ID) & Visual_ID != "") %>%  # Remove NA and empty values
  mutate(Present = 1) %>%  
  pivot_wider(names_from = Visual_ID, values_from = Present, values_fill = 0)

# Remove non-numeric columns
barcode_matrix <- barcode_pa %>% select(where(is.numeric))
morpho_matrix <- morpho_pa %>% select(where(is.numeric))

# Calculate dissimilarity matrices for both datasets
barcode_dist <- vegdist(barcode_matrix, method = "jaccard")  # Barcoding data
morpho_dist <- vegdist(morpho_matrix, method = "jaccard")    # Morphological data

# Convert dissimilarity matrices to data frames
barcode_df <- as.data.frame(as.matrix(barcode_dist))
morpho_df <- as.data.frame(as.matrix(morpho_dist))

# Add a 'method' column to distinguish between morphological and barcoding methods
barcode_df$method <- "Barcoding"
morpho_df$method <- "Morphological"

# Combine the two dataframes (one for barcoding and one for morphological data)
combined_df <- rbind(barcode_df, morpho_df)

# Add Shark_species information to the combined dataset
combined_df$Shark_Species <- rep(barcode_pa$Shark_species, 2)  # Assuming same samples for both methods

# Select only the numeric columns for the dissimilarity matrix
combined_df_numeric <- combined_df %>%
  select_if(is.numeric)

# Run PERMANOVA with method and Shark_Species as grouping factors
combined_perm <- adonis2(combined_df_numeric ~ method, data = combined_df, permutations = 999)
print(combined_perm)


####### Atlantic Sharpnose ########

#Now only by shark species (Atlantic Sharpnose) - this is FOO
# Filter the dataset for only Atlantic Sharpnose shark species
Chunk_per_AS <- Chunk_per %>% filter(Shark_species == "Atlantic Sharpnose")

# Convert BLAST_ID (DNA barcoding results) to presence-absence matrix
barcode_pa <- Chunk_per_AS %>%
  filter(!is.na(BLAST_ID) & BLAST_ID != "") %>%  # Remove NA and empty species names
  mutate(Present = 1) %>%  # Create presence indicator
  pivot_wider(names_from = BLAST_ID, values_from = Present, values_fill = 0)

# Convert Visual_ID (Morphological results) to presence-absence matrix
morpho_pa <- Chunk_per_AS %>%
  filter(!is.na(Visual_ID) & Visual_ID != "") %>%  # Remove NA and empty values
  mutate(Present = 1) %>%  
  pivot_wider(names_from = Visual_ID, values_from = Present, values_fill = 0)

# Remove non-numeric columns
barcode_matrix <- barcode_pa %>% select(where(is.numeric))
morpho_matrix <- morpho_pa %>% select(where(is.numeric))

# Calculate dissimilarity matrices for both datasets
barcode_dist <- vegdist(barcode_matrix, method = "jaccard")  # Barcoding data
morpho_dist <- vegdist(morpho_matrix, method = "jaccard")    # Morphological data

# Convert dissimilarity matrices to data frames
barcode_df <- as.data.frame(as.matrix(barcode_dist))
morpho_df <- as.data.frame(as.matrix(morpho_dist))

# Add a 'method' column to distinguish between morphological and barcoding methods
barcode_df$method <- "Barcoding"
morpho_df$method <- "Morphological"

# Combine the two dataframes (one for barcoding and one for morphological data)
combined_df <- rbind(barcode_df, morpho_df)

# Add Shark_species information to the combined dataset
combined_df$Shark_Species <- rep(barcode_pa$Shark_species, 2)  # Assuming same samples for both methods

# Select only the numeric columns for the dissimilarity matrix
combined_df_numeric <- combined_df %>%
  select_if(is.numeric)

# Run PERMANOVA with method and Shark_Species as grouping factors
combined_perm <- adonis2(combined_df_numeric ~ method, data = combined_df, permutations = 999)
print(combined_perm)
```